---
title: "p8105_hw3_cm3928"
author: "Clement Mugenzi"
date: "10/3/2019"
output: github_document
---



```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(skimr)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


# Question 1


```{r}
library(p8105.datasets)
data(instacart)
janitor::clean_names(instacart)
```

## A brief description of the dataset

The instacart dataset is comprised of **`r ncol(instacart)`** variables and **`r nrow(instacart)`** observations.

** I am going to group_by and summarize character variables.  

```{r}
doyou = 
  instacart %>% 
  group_by(product_name, department) %>% 
  summarise(n = n())
view(doyou)
```



**How many aisles are there, and which aisles are the most items ordered from?**

```{r}
num_aisle = 
  instacart %>% 
  count(aisle) %>% 
  summarise(n = n())
num_aisle
```


```{r}
items_ord = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(items_ordered = n()) %>% 
  arrange(-items_ordered)
view(items_ord)
```

```{r}
kable(items_ord, markdown)
```

There are 134 aisles and fresh vegetables is the most popular aisle with 150609 items ordered from it.



**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**


```{r}
plot = 
  items_ord %>%
  filter(items_ordered > 10000) %>% 
  ggplot(aes(x = items_ordered, y = aisle)) + 
  geom_point() +
  labs(
     x = "Number of items ordered") +
  scale_x_continuous(
    breaks = c(25000, 50000, 75000, 100000, 125000, 150000)
  )
plot
```


**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
tab = 
  instacart %>% 
  group_by(aisle, product_name) %>% 
  filter(min_rank(desc(order_id)) < 4) %>% 
  select(-c(product_id:days_since_prior_order), 
         -department_id, -aisle_id) %>% 
  summarise(order_id = n()) %>%
  filter(aisle == c("baking ingredients", "dog food care", 
                    "packaged vegetable fruits"))
view(tab)
```



**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r}
  instacart %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(order_dow, product_name) %>%
  summarise(
    mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) %>% 
knitr::kable(digits = 0)
```


























