---
title: "p8105_hw3_cm3928"
author: "Clement Mugenzi"
date: "10/3/2019"
output: github_document
---



```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(skimr)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "right"))
```


# Question 1


```{r}
library(p8105.datasets)
data(instacart)
janitor::clean_names(instacart)
```

## A brief description of the dataset

The instacart dataset is comprised of **`r ncol(instacart)`** variables and **`r nrow(instacart)`** observations.

I am going to group_by and summarize character variables.

**How many aisles are there, and which aisles are the most items ordered from?**

```{r}
num_aisle = 
  instacart %>% 
  count(aisle) %>% 
  summarise(n = n())
num_aisle
```


```{r}
items_ord = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(items_ordered = n()) %>% 
  arrange(-items_ordered)
```

```{r}
kable(head(items_ord), markdown)
```

There are 134 aisles and fresh vegetables is the most popular aisle with 150609 items ordered from it.



**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**


```{r}
  items_ord %>%
  filter(items_ordered > 10000) %>% 
  mutate(
    aisle = fct_reorder(aisle, items_ordered)) %>% 
  ggplot(aes(x = items_ordered, y = aisle)) + 
  geom_point(aes(color = aisle)) +
  labs(
    title = "Number of items ordered per aisle",
     x = "Number of items ordered") +
  scale_x_continuous(
    breaks = c(25000, 75000, 150000))
```


**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
  instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", 
                    "packaged vegetables fruits")) %>%
  select(-c(product_id:days_since_prior_order), 
         -department_id, -aisle_id, -department) %>% 
  group_by(aisle, product_name) %>% 
  summarise(order_id = n()) %>% 
  filter(min_rank(desc(order_id)) < 4) %>% 
  arrange(desc(order_id)) %>% 
  knitr::kable()
```



**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

```{r}
  instacart %>% 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(order_dow, product_name) %>%
  summarise(
    mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) %>% 
knitr::kable(digits = 1)
```


# Question 2



First, let's read in the dataset.

```{r}
library(p8105.datasets)
data(brfss_smart2010)
```

** a. Cleaning the dataset**

I renamed the **locationabbr** to **states.** and **locationdesc** to **counties.**

I used **mutate** to change the response variable to a factor variable so I can be able to arrange from **Poor** to **Excellent.**

```{r}
new_brfss_smart2010 = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(states = "locationabbr", counties = "locationdesc") %>% 
  filter(topic == "Overall Health", 
         response %in% c("Poor", "Fair","Good",
                       "Very good", "Excellent")) %>%
    mutate(response = factor(response, levels = c("Poor", "Fair","Good",
                       "Very good", "Excellent"))) %>% 
  arrange(response)
view(new_brfss_smart2010)
```



**b.**

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
  new_brfss_smart2010 %>%
  filter(year == "2002") %>%
  group_by(states) %>%
  distinct(counties) %>%
  summarise(
    location_02 = n()
  ) %>%
  filter(location_02 >= 7) %>% 
  knitr::kable()
```

The states that were observed at 7 or more locations in the year 2002 was:
**Connecticut, Florida, Massachusetts, North Carolina, New Jersey, and Pennsylvania.**

```{r}
  new_brfss_smart2010 %>%
  filter(year == "2010") %>%
  group_by(states) %>%
  distinct(counties) %>%
  summarise(
    location_10 = n()
  ) %>%
  filter(location_10 >= 7) %>% 
  knitr::kable()
```

States that were observed at 7 or more locations in the year 2010 were:
**California, Colorado, Florida, Massachusetts, Maryland, North Carolina, Nebraska, New Jersey, New York, Ohio, Pennsylvania, South Carolina, Texas, Washington.**


Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 


```{r}
spaghetti_df = 
  new_brfss_smart2010 %>%
  select(year, states, response, data_value) %>%
  filter(response == "Excellent") %>%
  group_by(year, states) %>% 
  summarise(
    mean_value = mean(data_value, na.rm = T))
```


```{r}
kable(spaghetti_df, markdown)
```

Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
spaghetti_plot = 
  spaghetti_df %>%
  ggplot(aes(x = year, y = mean_value, color = year)) +
  geom_line(aes(color = states))
spaghetti_plot
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.


```{r}
panel_plot = 
  new_brfss_smart2010 %>% 
  filter(response %in% c("Poor", "Very good", "Fair", "Good", "Excellent"), 
         year %in% c("2006", "2010"), states == "NY") %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = 0.5) +
  facet_grid(~year)
panel_plot
```



# Question 3


**Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).**

```{r}
accel_df = 
  read_csv("accel_data.csv") %>% 
  janitor::clean_names() %>% view()
```









































